An Algorithm For Spatial Mapping Using a Hexagon Tilegram, With Application to Australian Maps

### 2019-03-27 algorithmRjournal/algorithmRjournal.Rmd
Make plots more readable, need to add captions and figure numbers

```{r, fig.cap="lga Local Government Areas and derived centroids."}
end_cents <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude), data= centroids) + theme_void() + coord_equal()
end_cents

@@

hex_size <- 0.2

@@

```{r plot_grid}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r plot_shift}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r plot_buffer}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a", fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = full_grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75)
```

@@

Create a box filter

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

Create circle filter

@@

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)
```

Filter for angle within circle, relates hexagon location to focal point, Sydney.

        if (hex_grid$angle_minus[1] < hex_grid$angle_plus[1]) {
            hex_grid <- hex_grid %>%
                # create slice of 60 degrees from centroid
                filter(angle_minus < hex_angle & hex_angle < angle_plus)
        } else {
            hex_grid <- hex_grid %>%
                # create slice of 60 degrees from centroid
                filter(hex_angle < angle_plus | angle_minus > hex_angle)
        }

```{r}

ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

```{r}

# Choose first available point
        cent <- centroid1 %>% dplyr::rename(focal_point = points, focal_dist = focal_distance, focal_angle = angle)

        #update grid to show this centroid as assigned
        hex_grid[which(hex_grid$id == hex$hex_id),]$assigned <- TRUE

        centroid_allocation <- bind_rows(centroid_allocation, dplyr::bind_cols(cent, hex)) %>% as_tibble()

Using the angle between each polygon centroid, and it's the closest focal point, the subset of points is filtered again.
Of these possible points, only those within a specific amount of degrees plus and minus a specific angle range are kept. This angle begins at 30 degrees by default, and may increase if necessary.

@@

@@

```{r}
ggplot(hex_points_df) + 
  geom_polygon(aes(x = long, y = lat, group = LGA_CODE16)) +
  coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)
```

```{r fullhexmap}
# Create centroids set
centroids <- create_centroids(tas_lga, "LGA_CODE16")
# Create hexagon location grid
grid <- create_grid(centroids = centroids, hex_size = 0.25, buffer_dist = 1.2)
# Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(centroids = centroids,
hex_grid = grid,
sf_id = "LGA_CODE16",
hex_size = 0.25, # same size used in create_grid
hex_filter = 10,
focal_points = capital_cities,
width = 30, verbose = FALSE) # same column used in create_centroids
```

```{r}
h1 <- hex_allocated %>%
  fortify_hexagon(hex_size = 0.25) %>% 
  left_join(., tas_lga)

@@

end_hex <- ggplot() +
  geom_polygon(data = p1, aes(x=long, lat, group = group), alpha = 0.1) +
  geom_polygon(data = h1, aes(x=long, lat, group = LGA_CODE16)) + theme_void() + coord_equal()

gridExtra::grid.arrange(end_hex, end_cents)
```

## Applications of algorithm

<!--
Is  your  solution  always  working?  
Can  you  prove it? 
How does it work? 
-->

Examples of Melbourne, lga, Australia
At different levels, distance from original centroids, change in area

LGA of all Australia
SA2 for lga

<!--
Is your analysis about worst case or average case? 
What kind of theoretical and analytical support can you give for your proposed scheme/solution? 
-->

>>>>

```{r, fig.cap="lga Local Government Areas and derived centroids."}
end_cents <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data = centroids) + theme_void() + coord_equal()
end_cents

@@

hex_size <- 0.25

@@

```{r plot_grid}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r plot_shift}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data = centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r plot_buffer}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data = centroids, colour = "#b2df8a", fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = full_grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75)
```

@@

Create a box of possible hexagon locations areoung the centroid. The bottom of the box will not be square as the buffer has already removed unnecessary points from over the ocean.

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data = centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

The algorithm removes the outer corners of the square, creating a cricle of points, by only keeping points within a certain radius around the original centroid location.

@@

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data = centroid1, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)
```

The width parameter is now used. Of the remaining points, a slice is taken. This uses the angle from the closest capital city, to the current centroid. If it is not the closest to the capital there may have been points available that were taken by closer centroids. This allows the spatial relationship to be preserved, even when it gets allocated to a hexagon that is further from the city centre. This angle begins at 30 degrees by default, and may increase if necessary.

@@

        if (hex_grid$angle_minus[1] < hex_grid$angle_plus[1]) {
            hex_grid <- hex_grid %>%
                # create slice of 60 degrees from centroid
                filter(angle_minus < hex_angle & hex_angle < angle_plus)
        } else {
            hex_grid <- hex_grid %>%
                # create slice of 60 degrees from centroid
                filter(hex_angle < angle_plus | angle_minus > hex_angle)
        }
```

```{r}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75) + 
  geom_point(aes(x=longitude, y = latitude), data = centroid1, colour = "#b2df8a", size = 0.8) 
```

This area was the very closest to the Hobart central point ( coloured pink) provided in the capital cities data set, it gets a hexagon point (coloured blue) extremely close to it's centroid (coloured green).

```{r}
# Choose first available point
cent <- centroid1 %>% dplyr::rename(focal_point = points, focal_dist = focal_distance, focal_angle = angle)

# Update grid to show this centroid as assigned
hex_grid[which(hex_grid$id == hex$hex_id),]$assigned <- TRUE

# Add to table of allocated centroids
centroid_allocation <- bind_rows(centroid_allocation, dplyr::bind_cols(cent, hex)) %>% as_tibble()

@@

```{r}
ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_lga %>% filter(LGA_CODE16=="62810"), fill = "grey", size = 0.3, colour = "lightgrey") +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75) + 
  geom_point(aes(x=longitude1, y = latitude1), data = centroid1, colour = "#fb9a99", size = 1) + 
  geom_point(aes(x=longitude, y = latitude), data = centroid1, colour = "#b2df8a", size = 0.8) + 
  theme_void() +
  coord_equal()
```

The following code creates a map for all the Local Government Areas in Tasmania:

```{r fullhexmapcode, echo = TRUE}
# Create centroids set
centroids <- create_centroids(tas_lga, "LGA_CODE16")
# Create hexagon location grid
grid <- create_grid(centroids = centroids,
    hex_size = 0.25,
    buffer_dist = 1.2)
# Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "LGA_CODE16",
  # same column used in create_centroids
  hex_size = 0.25,
  # same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities %>% filter(points == "Hobart"),
  width = 30,
  verbose = FALSE
)
```

```{r, fig.height=2.5, fig.width = 5}
h1 <- hex_allocated %>%
  fortify_hexagon(hex_size = 0.25) %>% 
  left_join(., tas_lga)

@@

end_hex <- ggplot() +
  geom_polygon(data = p1, aes(x=long, lat, group = group), fill = "grey", size = 0.3, colour = "lightgrey") +
  geom_polygon(data = h1, aes(x=long, lat, group = LGA_CODE16)) + theme_void() + coord_equal()

gridExtra::grid.arrange(end_cents, end_hex, nrow = 1)
```

## Summary

It is possible to use alternative maps to communicate spatial distributions.
Current methods do not always work for Australia due to the large gaps between denseley populated capital cities.

## Code

This map algorithm is found in the `sugarbag` package, currently available on github: https://github.com/srkobakian/sugarbag

### 2019-03-27 algorithmRjournal/algorithmRjournal.tex
Make plots more readable, need to add captions and figure numbers

Create a box filter

@@

Filter for angle within circle, relates hexagon location to focal point,
Sydney.

Using the angle between each polygon centroid, and it's the closest
focal point, the subset of points is filtered again. Of these possible
points, only those within a specific amount of degrees plus and minus a
specific angle range are kept. This angle begins at 30 degrees by
default, and may increase if necessary.

@@

@@

@@

\hypertarget{applications-of-algorithm}{%
\subsection{Applications of algorithm}\label{applications-of-algorithm}}

Examples of Melbourne, lga, Australia At different levels, distance from
original centroids, change in area

LGA of all Australia SA2 for lga

>>>>

Create a box of possible hexagon locations areoung the centroid. The
bottom of the box will not be square as the buffer has already removed
unnecessary points from over the ocean.

@@

The width parameter is now used. Of the remaining points, a slice is
taken. This uses the angle from the closest capital city, to the current
centroid. If it is not the closest to the capital there may have been
points available that were taken by closer centroids. This allows the
spatial relationship to be preserved, even when it gets allocated to a
hexagon that is further from the city centre. This angle begins at 30
degrees by default, and may increase if necessary.

@@

This area was the very closest to the Hobart central point ( coloured
pink) provided in the capital cities data set, it gets a hexagon point
(coloured blue) extremely close to it's centroid (coloured green).

@@

The following code creates a map for all the Local Government Areas in
Tasmania:

\begin{Schunk}
\begin{Sinput}
# Create centroids set
centroids <- create_centroids(tas_lga, "LGA_CODE16")
# Create hexagon location grid
grid <- create_grid(centroids = centroids,
    hex_size = 0.25,
    buffer_dist = 1.2)
# Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "LGA_CODE16",
  # same column used in create_centroids
  hex_size = 0.25,
  # same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities %>% filter(points == "Hobart"),
  width = 30,
  verbose = FALSE
)
\end{Sinput}
\end{Schunk}

@@

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

It is possible to use alternative maps to communicate spatial
distributions. Current methods do not always work for Australia due to
the large gaps between denseley populated capital cities.

\hypertarget{code}{%
\subsection{Code}\label{code}}

This map algorithm is found in the \texttt{sugarbag} package, currently
available on github: \url{https://github.com/srkobakian/sugarbag}
